---
title: "lumber strength report"
author: "Yunfeng Yang"
date: "2023-06-07"
output: html_document
---

# Damage model 
The original model is 

\begin{equation}\label{eq: original model}
y^*  = y \cdot I\{c \cdot y > l\} + \alpha \cdot y \cdot I\{c \cdot y < l\}
\end{equation}

<!-- We may first consider a simplified version without the tuning parameter $c$. -->
<!-- \begin{equation}\label{eq: simplified model} -->
<!-- y^*  = y \cdot I\{ y > l\} + \alpha \cdot y \cdot I\{y < l\} -->
<!-- \end{equation} -->

# Smooth function

Note that we have the indicator function in the model we might consider the smooth function:
\begin{equation}\label{eq: trans function}
S(x;s) = \frac{1}{1 + \exp(-s \cdot x)}
\end{equation},
where $s$ is the smoothing hyper-parameter to control the smoothness. 

```{r}
smooth_ind <- function(x,s){
  return(1/(1+exp(-x*s)))
}
x <- seq(from = -10, to= 10, length.out = 500)
plot(x,ifelse(x>0,1,0),type = "l")
lines(x,sapply(x,smooth_ind, s = 1),type = "l",col = "red")
lines(x,sapply(x,smooth_ind, s = 10),type = "l",col = "blue")
lines(x,sapply(x,smooth_ind, s = 100),type = "l",col = "green")
legend("topleft", legend=c("Indicator", "s = 1", "s = 10", "s = 100"),
       col=c("black","red", "blue","green"), lty=1, cex=0.8)
```


Then the model becomes 
$$
y^*  = y \cdot \frac{1}{1 + \exp(-s\cdot(cy-l))} + \alpha \cdot y \cdot \frac{1}{1 + \exp(-s\cdot(l-cy))}
$$
# the plot 
```{r}
### new 

mu <- 48
sigma <- 19
#y <- rnorm(N,mean = mu, sd = sigma)
y <- seq(from = 0.1, to = 100, length = 500)
l <-  32
alpha <- 0.45
c <- 0.65
s <- 1
ystar <- y*smooth_ind(c*y-l,s )+ alpha*smooth_ind(l-c*y,s)

plot(y,ystar,type = 'l',lty = 1)
lines(y,y,lty = 2)
legend("topleft", legend = c("ystar","y"),lty=1:2)
```

We have $y \geq y^*$.

The lumber have three groups:

- Group 1: $y <l$, $y^* < l$, i.e., $y^* <y < l$. The lumber pieces are broken blow the proof loading.
- Group 2: $y >l$, $y^* < l$, i.e. $y^*<l<y$. The lumber pieces are broken during the proof loading process. This groups we only knows how many pieces. 
- Group 3: $y >l$, $y^* > l$, i.e. $l<y^*<y$. The lumber pieces survived in the proof-loading. And then we destruct them to test their strength.


# The PDF calculation 

Given $Y \sim N(\mu, \sigma^2)$, $Y^* = h(Y)$. Then the pdf of $Y^*$,
$$
f_{Y^*}(y^*) = f_{Y}(h^{-1}(y^*))|\frac{d}{dy^*}h^{-1}(y^*)|,
$$
where $f_Y()$ is the pdf of $Y$, i.e., normal. 


Following this, We need the numerical function of $h^{-1}(y^*)$, and its numerical gradient $\frac{d}{dy^*}h^{-1}(y^*)$. (The analytical form doesn't seem available.)
```{r}

#' The smooth function for the indicator function, which is also called "sigmoid function".
#' 
#' @param x, the variable 
#' @param s, the hyper parameter, higher s means closer the indicator function
#' @returns the smoothed value.  

smooth_ind <- function(x,s){
  return(1/(1+exp(-x*s)))
}



#' The damage model, smoothed version of 
#' "y* = y* I(c*x > l) + alpha * y * I(c*x < l)"
#' 
#' @param y, the original strength
#' @param alpha, the damage parameter
#' @param l, the proof loading level
#' @param c, the threshold parameter. Damage effects happen exceeding c.
#' @param s, the temperature parameter for s
#' @return, the weakened y*, y*<y.
dmg_model <- function(y,alpha,l,c,s){
  return(y*smooth_ind(c*y-l,s)+ alpha*y*smooth_ind(l-c*y,s))
}

#' The damage-inverse model, given a weakened ystar, find the original y
#' 
#' @param ystar, the weakened strength
#' @param alpha, the damage parameter
#' @param l, the proof loading level
#' @param c, the threshold parameter. Damage effects happen exceeding c.
#' @param s, the temperature parameter for s
#' @return, the original y, y>y*.
dmg_inverse <- function(ystar,alpha,l,c,s){
  uniroot((function (x) dmg_model(x,alpha,l,c,s) - ystar), lower = 0, upper = 1000)$root
}
# the abs gradient at damage-inverse
dmg_inverse_grad <- function(ystar,alpha,l,c,s){
  abs(numDeriv::grad(func = (function(x) dmg_inverse(x,alpha,l,c,s)),
                     x  = ystar))
}  
# dmg_inverse_grad(42,a lpha,l,c,s)
# ystar <- y*smooth_ind(c*y-l)+ alpha*smooth_ind(l-c*y)

dmglik <- function(ystar,alpha,l,c,s,mu,sigma){
  y <- dmg_inverse(ystar,alpha,l,c,s)
  dnorm(y, mean = mu, sd = sigma, log = TRUE)+ log(dmg_inverse_grad(ystar,alpha,l,c,s))
}
```


```{r}


## plot 
set.seed(8888)
# the orignal sample size 
N <- 300
mu <- 48
sigma <- 19
#y <- rnorm(N,mean = mu, sd = sigma)
l <-  32
# alpha <- 0.9
# c <- 0.8
alpha <- 0.45
c <- 0.65
s <- 1

## data 

y <- rnorm(N, mean = mu, sd = sigma )
y <- y[y>0]
y_obs <- list()
# group 1, y^* <y < l
y_obs$group1 <- y[which(y < l)]
# group 2, y^*<l<y
group23 <- y[which(y > l)]
group23_star <- dmg_model(group23,alpha,l,c,s)
y_obs$group2 <- length(group23[which(group23 > l &group23_star < l)])
y_obs$group3 <- group23_star[which(group23_star >l)]

negdmglik_model <- function(theta){
  mu <- theta[1]
  sigma <- theta[2]
  alpha <- theta[3]
  c <- theta[4]
  
  lik1 <- sum(dnorm(y_obs$group1,mean = mu, sd = sigma, log = TRUE))
  lik2 <- y_obs$group2 * log(
    pnorm(dmg_inverse(l,alpha,l,c,s), mean = mu, sd = sigma) -
      pnorm(l, mean = mu, sd = sigma)
  )
  #lik2 <- 0
  lik3 <- sum(sapply(y_obs$group3,dmglik,alpha,l,c,s,mu,sigma))
  return(-lik1-lik2-lik3)
  #return (-lik3)
}

theta0 <- c(mu,sigma,alpha,c)

```

```{r}
optimout <- optim(theta0,negdmglik_model,method = "L-BFGS-B",
                  lower = rep(0.1,4),upper = c(Inf,Inf,1,Inf))

# theta0 is true theta
rbind(theta0,optimout$par)
```



```{r}
optimCheck::optim_proj(optimout$par,
                       negdmglik_model,xrng = .5,
                       xnames = c("mu","sigma","alpha","c"))
```

<!-- # ```{r} -->
<!-- # ## plot  -->
<!-- # set.seed(8888) -->
<!-- # N <- 300 -->
<!-- # mu <- 48 -->
<!-- # sigma <- 19 -->
<!-- # #y <- rnorm(N,mean = mu, sd = sigma) -->
<!-- # l <-  32 -->
<!-- # alpha <- 0.45 -->
<!-- # c <- 0.65 -->
<!-- # s <- 1 -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # ## data  -->
<!-- #  -->
<!-- # y <- rnorm(N, mean = mu, sd = sigma ) -->
<!-- # y <- y[y>0] -->
<!-- # y_obs <- list() -->
<!-- # # group 1, y^* <y < l -->
<!-- # y_obs$group1 <- y[which(y < l)] -->
<!-- # # group 2, y^*<l<y -->
<!-- # group23 <- y[which(y > l)] -->
<!-- # group23_star <- dmg_model(group23,alpha,l,c,s) -->
<!-- # y_obs$group2 <- length(group23[which(group23 > l &group23_star < l)]) -->
<!-- # y_obs$group3 <- group23[which(group23_star >l)] -->
<!-- #  -->
<!-- # negdmglik_model <- function(theta){ -->
<!-- #   mu <- theta[1] -->
<!-- #   sigma <- theta[2] -->
<!-- #   alpha <- theta[3] -->
<!-- #   c <- theta[4] -->
<!-- #    -->
<!-- #   lik1 <- sum(dnorm(y_obs$group1,mean = mu, sd = sigma, log = TRUE)) -->
<!-- #   lik2 <- y_obs$group2 * log( -->
<!-- #     pnorm(dmg_inverse(l,alpha,l,c,s), mean = mu, sd = sigma) - -->
<!-- #       pnorm(l, mean = mu, sd = sigma) -->
<!-- #   ) -->
<!-- #   #lik2 <- 0 -->
<!-- #   lik3 <- sum(sapply(y_obs$group3,dmglik,alpha,l,c,s,mu,sigma)) -->
<!-- #   return(-lik1-lik2-lik3) -->
<!-- # } -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # theta0 <- c(mu,sigma,alpha,c) -->
<!-- # optimout <- optim(theta0,negdmglik_model,method = "L-BFGS-B", -->
<!-- #                   lower = rep(0.1,4),upper = c(Inf,Inf,1,Inf)) -->
<!-- #  -->
<!-- # optimCheck::optim_proj(optimout$par, -->
<!-- #                        negdmglik_model,xrng = .5, -->
<!-- #                        xnames = c("mu","sigma","alpha","c")) -->
<!-- # ``` -->

<!-- ## Comparison of the likelihood in group2 and group3 -->

<!-- ```{r} -->
<!-- alpha_seq <- seq(from = 0.01, to = 0.99, by = 0.01) -->

<!-- ## group 2 likelihood  -->
<!-- negdmglik_alpha_g2  <- function(alpha){ -->
<!--   mu <- mu -->
<!--   sigma <- sigma -->
<!--   c <- c -->
<!--     lik2 <- y_obs$group2 * log( -->
<!--     pnorm(dmg_inverse(l,alpha,l,c,s), mean = mu, sd = sigma) - -->
<!--       pnorm(l, mean = mu, sd = sigma) -->
<!--   ) -->

<!--   return (-lik2) -->
<!-- } -->
<!-- neglikg2 <- sapply(alpha_seq,negdmglik_alpha_g2) -->

<!-- ## group 3 likelihood  -->
<!-- negdmglik_alpha_g3  <- function(alpha){ -->
<!--   mu <- mu -->
<!--   sigma <- sigma -->
<!--   c <- c -->


<!--   #lik2 <- 0 -->
<!--   lik3 <- sum(sapply(y_obs$group3,dmglik,alpha,l,c,s,mu,sigma)) -->
<!--   # return(-lik1-lik2-lik3) -->
<!--   return (-lik3) -->
<!-- } -->
<!-- neglikg3 <- sapply(alpha_seq,negdmglik_alpha_g3) -->

<!-- ``` -->

<!-- The plot of negative likelihood of alpha in group 2. -->
<!-- ```{r} -->
<!-- plot(alpha_seq,neglikg2, type = "l") -->
<!-- ``` -->

<!-- The plot of negative likelihood of alpha in group 3. -->

<!-- ```{r} -->
<!-- plot(alpha_seq,neglikg3, type = "l") -->
<!-- ``` -->



<!-- ```{r} -->
<!-- negdmglik_alpha  <- function(alpha){ -->
<!--   mu <- mu -->
<!--   sigma <- sigma -->
<!--   c <- c -->

<!--   lik1 <- sum(dnorm(y_obs$group1,mean = mu, sd = sigma, log = TRUE)) -->
<!--   lik2 <- y_obs$group2 * log( -->
<!--     pnorm(dmg_inverse(l,alpha,l,c,s), mean = mu, sd = sigma) - -->
<!--       pnorm(l, mean = mu, sd = sigma) -->
<!--   ) -->
<!--   #lik2 <- 0 -->
<!--   lik3 <- sum(sapply(y_obs$group3,dmglik,alpha,l,c,s,mu,sigma)) -->
<!--   # return(-lik1-lik2-lik3) -->
<!--   return (-lik2-lik3) -->
<!-- } -->

<!-- alpha_seq <- seq(from = 0.01, to = 0.99, by = 0.01) -->
<!-- negloglik <- sapply(alpha_seq,negdmglik_alpha) -->

<!-- ``` -->

<!-- The plot of negatgive likelihood in group 2 and group 3(i.e., likelihood in group 2 + likelihood in group 3). It seems that they are not balanced. And the group 2 dominates the likelihood.  -->

<!-- ```{r} -->
<!-- plot(alpha_seq,negloglik, type = "l") -->

<!-- ``` -->

# Real data analysis 

In this section, we apply the damage model to the real dataset. 

I also used the logit and expit transformation to map the damage effect parameter $\alpha$ (0,1) to R for convenience transformation. 

Further, the total sample size of the real dataset is N = 300.
```{r}
bending <- read.csv("bending-pl.csv", header = T)

bending[,1] <- bending[,1]/1000*6.895
l <- 4500/1000*6.895
# bending_mar <- bending[1:195,]
# bending_dmg <- bending[195:341,]

R100_data <- bending[bending[,2] == "R100",1]
R20_data <- bending[bending[,2] == "R20",1]
R20R100_data <- bending[bending[,2] == "R20R100",1]

# 97 pieces breaking in the loading process.
g2 <- 97
mu <- 48
sigma <- 19

s <- 1
```

```{r}

#' The smooth function for the indicator function, which is also called "sigmoid function".
#' 
#' @param x, the variable 
#' @param s, the hyper parameter, higher s means closer the indicator function
#' @returns the smoothed value.  

smooth_ind <- function(x,s){
  return(1/(1+exp(-x*s)))
}



#' The damage model, smoothed version of 
#' "y* = y* I(c*x > l) + alpha * y * I(c*x < l)"
#' 
#' @param y, the original strength
#' @param alpha, the damage parameter
#' @param l, the proof loading level
#' @param c, the threshold parameter. Damage effects happen exceeding c.
#' @param s, the temperature parameter for s
#' @return, the weakened y*, y*<y.
dmg_model <- function(y,alpha,l,c,s){
  return(y*smooth_ind(c*y-l,s)+ alpha*y*smooth_ind(l-c*y,s))
}

#' The damage-inverse model, given a weakened ystar, find the original y
#' 
#' @param ystar, the weakened strength
#' @param alpha, the damage parameter
#' @param l, the proof loading level
#' @param c, the threshold parameter. Damage effects happen exceeding c.
#' @param s, the temperature parameter for s
#' @return, the original y, y>y*.
dmg_inverse <- function(ystar,alpha,l,c,s){
  uniroot((function (x) dmg_model(x,alpha,l,c,s) - ystar), lower = 0, upper = 1000)$root
}
# the abs gradient at damage-inverse
dmg_inverse_grad <- function(ystar,alpha,l,c,s){
  abs(numDeriv::grad(func = (function(x) dmg_inverse(x,alpha,l,c,s)),
                     x  = ystar))
}  
# dmg_inverse_grad(42,a lpha,l,c,s)
# ystar <- y*smooth_ind(c*y-l)+ alpha*smooth_ind(l-c*y)

dmglik <- function(ystar,alpha,l,c,s,mu,sigma){
  y <- dmg_inverse(ystar,alpha,l,c,s)
  dnorm(y, mean = mu, sd = sigma, log = TRUE)+ log(dmg_inverse_grad(ystar,alpha,l,c,s))
}

logit <- function(x){
  return(log(x/(1-x)))
}

expit <- function(x){
  1/(1+exp(-x))
}

negdmglik_model <- function(theta){
  mu <- theta[1]
  sigma <- theta[2]
  # alpha <- theta[3]
  alpha <- expit(theta[3])
  c <- theta[4]
  
  lik1 <- sum(dnorm(R20_data,mean = mu, sd = sigma, log = TRUE))
  lik2 <- g2 * log(
    pnorm(dmg_inverse(l,alpha,l,c,s), mean = mu, sd = sigma) -
      pnorm(l, mean = mu, sd = sigma)
  )
  #lik2 <- 0
  lik3 <- sum(sapply(R20R100_data,dmglik,alpha,l,c,s,mu,sigma))
  lik4 <- sum(dnorm(R100_data,mean = mu, sd = sigma, log = TRUE))
  return(-lik1-lik2-lik3-lik4)
}

negdmglik_model_grad <- function(theta){
  numDeriv::grad(negdmglik_model,theta_test)
}
theta_test <- c(mu,sigma,logit(0.5),.6)




theta_test <- c(47.9985218, 18.9804222,logit(0.4548490),  0.6515348)
negdmglik_model(theta_test)
optimout <- optim(par = theta_test,
                  fn = negdmglik_model,
                  gr = negdmglik_model_grad,
                  method = "BFGS")
optimout$par

expit(optimout$par[3])


```


```{r}
s <- 1
negdmglik_model_unconstraint <- function(theta){
  mu <- theta[1]
  sigma <- theta[2]
  # alpha <- theta[3]
  alpha <- theta[3]
  c <- theta[4]
  
  lik1 <- sum(dnorm(R20_data,mean = mu, sd = sigma, log = TRUE))
  lik2 <- g2 * log(
    pnorm(dmg_inverse(l,alpha,l,c,s), mean = mu, sd = sigma) -
      pnorm(l, mean = mu, sd = sigma)
  )
  #lik2 <- 0
  lik3 <- sum(sapply(R20R100_data,dmglik,alpha,l,c,s,mu,sigma))
  lik4 <- sum(dnorm(R100_data,mean = mu, sd = sigma, log = TRUE))
  return(-lik1-lik2-lik3-lik4)
}

theta_opt <- optimout$par
theta_opt[3] <- expit(theta_opt[3])
optimCheck::optim_proj(theta_opt,
                       negdmglik_model_unconstraint ,xrng = .5,
                       xnames = c("mu","sigma","alpha","c"))
```

The plot shows that alpha does not seem to reach the mode.


## Real data analysis: control the other three parameter and only optimize alpha

```{r}
## the only plot for alpha

negdmglik_alpha <- function(alpha){
  mu <- 47.9985218
  sigma <- 18.9804222
  # alpha <- theta[3]
  c <- 0.6515348

  lik1 <- sum(dnorm(R20_data,mean = mu, sd = sigma, log = TRUE))
  lik2 <- g2 * log(
    pnorm(dmg_inverse(l,alpha,l,c,s), mean = mu, sd = sigma) -
      pnorm(l, mean = mu, sd = sigma)
  )
  #lik2 <- 0
  lik3 <- sum(sapply(R20R100_data,dmglik,alpha,l,c,s,mu,sigma))
  lik4 <- sum(dnorm(R100_data,mean = mu, sd = sigma, log = TRUE))
  return(-lik1-lik2-lik3-lik4)
}

s <- 1
alpha_opt <- optimize(negdmglik_alpha,c(.01,.99))


alpha_seq <- seq(from = 0.01, to = 0.5, length =200)
plot(alpha_seq, sapply(alpha_seq,negdmglik_alpha),type = "l", ylab = "neglik")
```

```{r}
alpha_opt$minimum
```

Although the likelihood not very smooth, but the plot does indicates that the mode is around 0.24.


# real data test

As all group3 has ystar (i.e., the damaged pieces), I tried to transform them back to the original y. 

```{r}
test_data <- cbind(R20R100_data,
                   alpha45 = sapply(R20R100_data, dmg_inverse,0.45,l,0.65,s),
                   alpha25 = sapply(R20R100_data, dmg_inverse,0.25,l,0.65,s))
test_data 
```

It seems that the result from $\alpha = 0.45$ and $\alpha = 0.25$ gives very similar y (the undamaged strength).


```{r}
alpha <- 0.25
c <- 0.65
y <- seq(from  = 0.01, to = 100, length = 200)
ystar <- dmg_model(y,alpha,l,c,s =1)
plot(y,ystar,type = 'l',lty = 1)
lines(y,dmg_model(y,alpha = 0.45,l,c,s =1),lty = 2)
legend("topleft", legend = c("ystar alpha = 0.45","ystar alpha = 0.25"),lty=1:2)
#abline(v = l)
abline(v = l/c)
```


<!-- # the loglik vs alpha  -->

<!-- ```{r} -->
<!-- negdmglik_alpha <- function(alpha){ -->
<!--   mu <- 47.9985218 -->
<!--   sigma <- 18.9804222 -->
<!--   # alpha <- theta[3] -->
<!--   c <- 0.6515348 -->

<!--   lik1 <- sum(dnorm(R20_data,mean = mu, sd = sigma, log = TRUE)) -->
<!--   lik2 <- g2 * log( -->
<!--     pnorm(dmg_inverse(l,alpha,l,c,s), mean = mu, sd = sigma) - -->
<!--       pnorm(l, mean = mu, sd = sigma) -->
<!--   ) -->
<!--   #lik2 <- 0 -->
<!--   lik3 <- sum(sapply(R20R100_data,dmglik,alpha,l,c,s,mu,sigma)) -->
<!--   lik4 <- sum(dnorm(R100_data,mean = mu, sd = sigma, log = TRUE)) -->
<!--   return(-lik1-lik2-lik3-lik4) -->
<!-- } -->


<!-- alpha <- seq(from = 0.01, to = 0.99, by = 0.01) -->
<!-- plot(alpha, sapply(alpha,negdmglik_alpha),type = "l", ylab = "neglik") -->
<!-- ``` -->



<!-- # check the likelihood of the damage model -->

<!-- In this section, I suppose all lumber samples are damaged. If the likelihood has the mode around the true value, it can indicate the likelihood is correct. -->

<!-- ```{r} -->


<!-- N <- 1000 -->
<!-- ## suppose all observations are damaged  -->
<!-- set.seed(8888) -->
<!-- y <- rnorm(N, mean = mu, sd = sigma ) -->
<!-- y <- y[y>0] -->
<!-- ystar <- dmg_model(y,alpha,l,c,s) -->

<!-- negg3 <- function(theta){ -->
<!--   mu <- theta[1] -->
<!--   sigma <- theta[2] -->
<!--   alpha <- theta[3] -->
<!--   c <- theta[4] -->
<!--   lik3 <- sum(sapply(ystar,dmglik,alpha,l,c,s,mu,sigma)) -->
<!--   return (-lik3) -->
<!-- } -->

<!-- negg3(theta0) -->

<!-- optimCheck::optim_proj(theta0, -->
<!--                        negg3, -->
<!--                        xrng = .5, -->
<!--                        xnames = c("mu","sigma","alpha","c")) -->



<!-- ``` -->

<!-- The likelihood has the mode around the true value (red lines). The likelihood of the damage model should be correct. -->